---
title: 'HW 4: Lecture 5'
author: 'Name: Jiacheng Wang'
output: 
  html_document:
    toc: TRUE
number_sections: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---

# Question 1

Recall the histogram for cut categories from the diamonds data set:

```{r, message=FALSE, warning=FALSE}
library(plotly)
library(tidyverse)

diamonds %>%
  plot_ly( x = ~cut, showlegend = FALSE ) %>%
  add_histogram(hoverinfo = "x+y") %>%
  dplyr::group_by( cut ) %>%
  summarise(n = n()) %>%
  add_text(
    text = ~scales::comma(n), y = ~n,
    textposition = "top middle",
    cliponaxis = FALSE,
    hoverinfo = "none"
  )
```

Please add to it the following changes:

1. make the bar chart a stacked bar chart: so that the x-axis values are still according to the variable cut, but the internal division of each bar is according to the `clarity` variable.
2. change the colors of the bar segments to be in accordance with the `"Blues"` palette.
3. make sure to have the tooltip content for each bar segment as it is in the below plot (“The cut is … such diamonds!”)
4. set the color of the font in the tooltip to red (hint: `hoverlabel`).
5. set the background of the tooltip to yellow.
6. set the background color of the entire plot to `"pink"`.
7. set the background color of the plotting area to `"lightgray"`.
8. set the borders of the bars to black and thickness (`span`) of 1.
9. set the color of the x-axis labels to `"green"`.
10. set the title of the x-axis to `"Cut"` (instead of “cut”).
11. set the color of the x-axis title to `"darkred"` and its font to `"Times New Roman"`.
12. change the color of the y-axis labels to `"darkorange"` (be careful not to change other things with it).
13. remove the y-axis title.
14. set the left margin of the plot to 25px.
15. change the color of the text values to `"red"`. No tooltip is associated with these values.
16. add the title `"Ugly, ugly plot"` in the `"Old Standard TT"` font at the top left corner.

In addition, make sure that the data that `plot_ly` utilizes is the minimum data necessary for the plot.

**Notes:**

1. The title of the sample plot is `"Pretty ugly plot"`, which I applied here instead of `"Ugly, ugly plot"`.
2. The color of the x-axis labels is required to be green, which I applied here instead of the color used in the sample plot.


```{r}
total_counts <- diamonds %>%
  group_by(cut) %>%
  summarise(total = n())

diamonds_summary <- diamonds %>%
  count(cut, clarity)

p <- diamonds %>%
  count(cut, clarity) %>%  # Count the number of diamonds for each combination of cut and clarity
  plot_ly() %>%
  add_bars(
    x = ~cut, 
    y = ~n, 
    color = ~clarity,
    colors = "Blues",
    hoverinfo = 'text',
    text = ~paste("The cut is", cut, "\nthe clarity", clarity, "\nand there are", n, "such diamonds!"),
    marker = list(line = list(color = 'black', width = 1)), # Set the borders of the bars to black with a thickness of 1
    showlegend = FALSE,
    textposition = "none"
    ) %>%
  layout(
    barmode = 'stack',
    hoverlabel = list(bgcolor = "yellow", font = list(color = "red")), # Set tooltip text color to red and background to yellow
    paper_bgcolor = "pink", # Set the background color of the entire plot to pink
    plot_bgcolor = "lightgray", # Set the background color of the plotting area to lightgray
    xaxis = list(
      tickfont = list(color = "green"), # Set the color of the x-axis labels to green
      title = "Cut", # Set the title of the x-axis to "Cut"
      titlefont = list(color = "darkred", family = "Times New Roman") # Set the color and font of the x-axis title
      ),
    yaxis = list(tickfont = list(color = "darkorange"), title = list(text = "")), # Set the color of the y-axis labels to darkorange and remove the y-axis title
    margin = list(l = 25), # Set the left margin of the plot to 25px
    title = list(text = "Pretty ugly plot", font = list(family = "Old Standard TT"), x = 0, y = 1) # Add the title "Ugly, ugly plot" in the "Old Standard TT" font at the top left corner
    )

# Add the text annotations for total counts
for (i in 1:nrow(total_counts)) {
  p <- p %>% add_text(
    data = total_counts[i, ],
    x = ~cut,
    y = ~total,
    text = ~total,
    textposition = 'top middle',
    hoverinfo = 'none',
    showlegend = FALSE,
    textfont = list(color = "red") 
  )
}

p
```





