---
title: "HW5"
author: "Owen Gallagher, Nikki Lin, Jiacheng Wang"
date: "2024-03-13"
output: 
  ioslides_presentation:
    widescreen: yes
    css: styles.css
runtime: shiny
---

```{r setup, include=FALSE, echo=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(plotly)
library(countrycode)


#load in world data
world.data <- read.csv('world-data-2023.csv')

#load in happiness data
happy.data <- read.csv('WHR2023.csv')

#clean happy.data
happy.data <- happy.data %>%
  select(c(Country.name, Ladder.score, Logged.GDP.per.capita,Social.support, 
           Healthy.life.expectancy, Freedom.to.make.life.choices, Generosity,
           Perceptions.of.corruption)) %>%
  rename(c(happiness.score = Ladder.score, Country = Country.name))

#Change some country names so they join properly
old.name <- c('Republic of Ireland', 'Czech Republic', 'Republic of the Congo', "Turkey", 'The Gambia', 
              'Democratic Republic of the Congo')
new.name <- c('Ireland', 'Czechia','Congo (Brazzaville)',
              'Turkiye','Gambia','Congo (Kinshasa)')

#Clean the world data by dropping certain rows and 
world.data <- world.data %>%
                select(-c(Capital.Major.City, Largest.city, Abbreviation, Currency.Code,Calling.Code)) %>%
                mutate(Country.join = replace(Country, Country %in% old.name, new.name)) %>%
                rename(Density.P.Km2 = Density..P.Km2., Agricultural.Land.Percent = Agricultural.Land....,
                       Land.Area.Km2 = Land.Area.Km2., CPI.Change = CPI.Change....,Forested.Area.Percent = Forested.Area....,
                       Gross.Primary.Education.Enrollment.Percent = Gross.primary.education.enrollment...., Gross.Tertiary.Education.Enrollment.Percent = Gross.tertiary.education.enrollment....,
                       Population.Labor.force.participation.Percent = Population..Labor.force.participation....,
                       Tax.Revenue.Percent = Tax.revenue....)


### Remove Percent Signs
remove.char <- function(x){
  
  #create vector to remove percent, comma, and dollar sign
    vec <- gsub('%','',x)
    vec <- gsub(',','',vec)
    vec <- gsub('\\$','',vec)
  
    #make numeric
    vec <- as.numeric(vec)
    
    #return vector
    return(vec)
}

world.data$Density.P.Km2 <- remove.char(world.data$Density.P.Km2)
world.data$Agricultural.Land.Percent <- remove.char(world.data$Agricultural.Land.Percent)
world.data$Land.Area.Km2 <- remove.char(world.data$Land.Area.Km2)
world.data$Armed.Forces.size <- remove.char(world.data$Armed.Forces.size)
world.data$Co2.Emissions <- remove.char(world.data$Co2.Emissions)
world.data$CPI <- remove.char(world.data$CPI)
world.data$CPI.Change <- remove.char(world.data$CPI.Change)
world.data$Forested.Area.Percent <- remove.char(world.data$Forested.Area.Percent)
world.data$Gasoline.Price <- remove.char(world.data$Gasoline.Price)
world.data$GDP <- remove.char(world.data$GDP)
world.data$Gross.Primary.Education.Enrollment.Percent <- remove.char(world.data$Gross.Primary.Education.Enrollment.Percent)
world.data$Gross.Tertiary.Education.Enrollment.Percent <- remove.char(world.data$Gross.Tertiary.Education.Enrollment.Percent)
world.data$Minimum.wage <- remove.char(world.data$Minimum.wage)
world.data$Out.of.pocket.health.expenditure <- remove.char(world.data$Out.of.pocket.health.expenditure)
world.data$Population <- remove.char(world.data$Population)
world.data$Population.Labor.force.participation.Percent <- remove.char(world.data$Population.Labor.force.participation.Percent)
world.data$Tax.Revenue.Percent <- remove.char(world.data$Tax.Revenue.Percent)
world.data$Total.tax.rate <- remove.char(world.data$Total.tax.rate)
world.data$Unemployment.rate <- remove.char(world.data$Unemployment.rate)
world.data$Urban_population <- remove.char(world.data$Urban_population)


# Join the two datasets
final.data <- happy.data %>%
                left_join(world.data, by = join_by(Country == Country.join)) %>%
                select(-Country.y) %>%
                dplyr::filter(!(Country %in% c('Kosovo', 'Hong Kong S.A.R. of China', 'State of Palestine', 'Taiwan Province of China', 'State of Palestine'))) %>%
                mutate(iso.code = countrycode::countrycode(Country, origin = 'country.name', destination = 'iso3c'))

# Write the text for the plot into the dataset as a variable
# This will be called in the plot later
final.data$hover <- with(final.data, paste(Country, '<br>', 'Population (in millions):', round(Population / 1000000,2), '<br>', 
                          'Life.expectancy: ', Life.expectancy, '<br>', 'GDP Per Cap: $', round(exp(Logged.GDP.per.capita),0)))

```


```{r echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
library(dplyr)
library(plotly)
library(shiny)
library(markdown)
library(shinythemes)
library(readr)
library(stargazer)
library(tidyverse)
library(leaps)
library(HH)
library(car)
library(reshape2)
library(broom)
library(countrycode)
library(plotly)
library(patchwork)
library(stargazer)
library(tidyverse)
library(plotly)
library(countrycode)

```


## Guiding Question

- What factors in a country lead to happiness? 
- Our group decided to look at factors such as 
  - Geography
  - Economics 
  - Health
  
## The Data

- The happiness data was pulled from the World Happiness Report 2023 (Source)[https://worldhappiness.report/data/]
- Additional country statistics were merged to the dataset from the following (Kaggle dataset)[https://www.kaggle.com/datasets/nelgiriyewithana/countries-of-the-world-2023]
- Finally the `countrycode` package was used to import the continent and the iso country code
- The dataset contains 133 countries
  - Kosovo, Hong Kong, State of Palestine, Taiwan, and Palestine were dropped because from the dataset because the country statistics dataset did not have values for them
  - The World Happiness Report did not survey every country in the world, so we **do not** have all countries



## Map of Happiness

```{r, out.width='100%'}
final.data <- read.csv('HW5.data.csv')

plot_ly(data = final.data, type = 'choropleth', locations = ~iso.code, z = ~happiness.score, 
        text = ~hover, colors = 'Spectral') %>%
  colorbar(title = 'Happiness Score') %>%
  layout(title = '2023 World Happiness Score for 133 Countries')
```

- The Western World appears to have the most happiness
- In particular the Nordic Countries (Finland, Norway, Sweden, and Denmark) have the most happiness
- African countries are the most unhappy, followed by countries in the middle east
- Surprisingly Ukraine is not incredibly unhappy



## Map of Happiness
```{r, echo=FALSE, warning=FALSE, message=FALSE}
hw5 <- read.csv("HW5.data.csv")

psydata <- hw5 %>% 
  dplyr::select(Country, happiness.score, Social.support, Freedom.to.make.life.choices, Perceptions.of.corruption) 

psydata2 <- psydata %>% 
    pivot_longer(c("Social.support", "Freedom.to.make.life.choices", "Perceptions.of.corruption"), names_to = "Predictor", values_to = "Score")

psydata2$Continent <- countrycode(sourcevar = psydata2[["Country"]], origin = "country.name",
                            destination = "continent")

hw5 <- read.csv("HW5.data.csv")

psydata <- hw5 %>% 
  dplyr::select(Country, happiness.score, Social.support, Freedom.to.make.life.choices, Perceptions.of.corruption) 

psydata2 <- psydata %>% 
    pivot_longer(c("Social.support", "Freedom.to.make.life.choices", "Perceptions.of.corruption"), names_to = "Predictor", values_to = "Score")

psydata2$Continent <- countrycode(sourcevar = psydata2[["Country"]], origin = "country.name",
                            destination = "continent")

psy_pred_model2 <- lm(happiness.score ~ Social.support + Freedom.to.make.life.choices + Perceptions.of.corruption, data = hw5)

plot1 <- ggplot(psydata2, aes(Score, happiness.score, color = Continent)) + 
  geom_point(aes(ids = Country), alpha = 0.5) +
  theme(legend.title = element_blank()) +
  facet_wrap(~Predictor, labeller = labeller(Predictor = 
    c("Freedom.to.make.life.choices" = "Life Choice Freedom",
      "Perceptions.of.corruption" = "Corruption Perception",
      "Social.support" = "Social Support")
  )) +
  geom_smooth(method='lm', se = F, alpha = 0.5) + 
  theme_bw() +
  ylab("Happiness Score")

plotly1 <- ggplotly(plot1)
plotly1

plot2 <- ggplot(psydata2, aes(x = Score, y = happiness.score, color = Predictor)) + 
  geom_smooth(method='lm', aes(fill = Predictor), se = TRUE, alpha = 0.5) +
  scale_fill_manual(values=c(Freedom.to.make.life.choices="plum3",
                             Perceptions.of.corruption="khaki1",
                             Social.support= "mistyrose")) +
    scale_color_manual(values=c(Freedom.to.make.life.choices="mediumpurple4",
                             Perceptions.of.corruption="orange2",
                             Social.support="palevioletred3")) +
  theme_bw() +
  ylab("Happiness Score")

plotly2 <- ggplotly(plot2, tooltip = c("text"))
plotly2

subplot(plotly2, plotly1, nrows = 2, margin = 0.065, heights = c(0.5, 0.5)) %>% 
  layout(legend = list(title = " ", font = list(size = 10)), 
         title = "Psychological Predictors of Happiness",
         annotations = list(
           list(x = -0.08 , y = 0.5, text = "Happiness Score",
                     font = list(size = 12),
                     textangle = 270,
                     showarrow = F, xref='paper', yref='paper', size=48),
           list(x = 0.5 , y = -0.10, text = "Predicter Score",
                     font = list(size = 12),
                     textangle = 0,
                     showarrow = F, xref='paper', yref='paper', size=48))
  )
```


## {data-background="#fff" .no-title-slide }
```{r shiny, echo=FALSE, warning=FALSE, message=FALSE}
# Data preparation
hw_data <- read.csv("HW5.data.csv", stringsAsFactors = TRUE)
countries_all <- read.csv("regions.csv", stringsAsFactors = TRUE)
combined_data <- left_join(hw_data, countries_all, by = "Country")
write.csv(combined_data, "HW5_with_regions.csv")
happiness2023 <- read.csv("HW5_with_regions.csv", stringsAsFactors = TRUE)
colnames <- colnames(happiness2023) %>% .[-c(1, 2, 3)] %>% .[-length(.)]

# Define UI 
ui <- navbarPage(
  theme = shinytheme("journal"),
  title = "World Happiness",
  tabPanel("Plot",
           sidebarPanel(
             width = 3,
             helpText("Choose variables you want to see"),
             uiOutput("variable_1"),
             uiOutput("variable_2"),
             helpText("Click while pressing `ctrl / shift` to choose multiple countries"),
             uiOutput("country"),
             hr(),
             span("Data source:", 
                  tags$a("WHR 2023",
                         href = "http://worldhappiness.report"))),
           mainPanel(
             plotlyOutput("scatterplot", height = 500))),
  
  tabPanel("About",
           includeMarkdown("About.md")))
      

# Define server 
server <- function(input, output, session) {
  
  output$variable_1 <- renderUI({
    selectInput("variable_1", "Y-variable:", 
                choices = colnames, selected = "happiness.score")})
  
  output$variable_2 <- renderUI({
    selectInput("variable_2", "X-variable:", 
                choices = colnames, selected = "happiness.score")})

  data_2023 <- reactiveValues(data = happiness2023)
  
  filtered_data <- reactive({
    data_2023$data %>% 
      filter(Country %in% c(input$country)) %>% 
      arrange(Country)
    })
  
  countries <- reactive({ 
    df_small <- 
      happiness2023 %>% 
      select(Country) %>% 
      droplevels()
    c("Select All", levels(df_small$Country))
    })
  
  output$country <- renderUI ({
    selectInput("country", "Countries:", 
                multiple = TRUE,
                choices = countries(), 
                selected = "Select All",
                selectize = FALSE,
                size = 10)
  })

  observe({
    if ("Select All" %in% input$country) {
      # choose all the choices _except_ "Select All"
      selected_choices <- setdiff(countries(), "Select All")
      updateSelectInput(session, "country", selected = selected_choices)
    }
  })

  output$scatterplot <- renderPlotly({
    if (is.null(data_2023$data)) return()
    
    p <- ggplot(filtered_data()) +
      geom_point(aes_string(x = input$variable_2, y = input$variable_1, 
                            colour = "Region", label = "Country"), size = 3) +
      ggtitle(paste0(input$variable_1, " vs. ", input$variable_2)) +
      theme_bw() +
      theme(legend.position = "bottom")
    
    ggplotly(p, tooltip = c("label", "x", "y"), height = 500) 
  })
}

# Run the application 
shinyApp(ui = ui, server = server)

```



