---
title: "Analyzing the Worlds Happiness"
author: "Owen Gallagher, Jia Qi Lin, Jiacheng Wang"
date: "`r Sys.Date()`"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(plotly)
library(countrycode)

#load in world data
world.data <- read.csv('world-data-2023.csv')

#load in happiness data
happy.data <- read.csv('WHR2023.csv')

#clean happy.data
happy.data <- happy.data %>%
                select(c(Country.name, Ladder.score, Logged.GDP.per.capita,Social.support, Healthy.life.expectancy, Freedom.to.make.life.choices,
                         Generosity,Perceptions.of.corruption)) %>%
                rename(c(happiness.score = Ladder.score, Country = Country.name))

#Change some country names so they join properly
old.name <- c('Republic of Ireland', 'Czech Republic', 'Republic of the Congo', "Turkey", 'The Gambia', 
              'Democratic Republic of the Congo')
new.name <- c('Ireland', 'Czechia','Congo (Brazzaville)',
              'Turkiye','Gambia','Congo (Kinshasa)')

#Clean the world data by dropping certain rows and 
world.data <- world.data %>%
                select(-c(Capital.Major.City, Largest.city, Abbreviation, Currency.Code,Calling.Code)) %>%
                mutate(Country.join = replace(Country, Country %in% old.name, new.name)) %>%
                rename(Density.P.Km2 = Density..P.Km2., Agricultural.Land.Percent = Agricultural.Land....,
                       Land.Area.Km2 = Land.Area.Km2., CPI.Change = CPI.Change....,Forested.Area.Percent = Forested.Area....,
                       Gross.Primary.Education.Enrollment.Percent = Gross.primary.education.enrollment...., Gross.Tertiary.Education.Enrollment.Percent = Gross.tertiary.education.enrollment....,
                       Population.Labor.force.participation.Percent = Population..Labor.force.participation....,
                       Tax.Revenue.Percent = Tax.revenue....)


### Remove Percent Signs
remove.char <- function(x){
  
  #create vector to remove percent, comma, and dollar sign
    vec <- gsub('%','',x)
    vec <- gsub(',','',vec)
    vec <- gsub('\\$','',vec)
  
    #make numeric
    vec <- as.numeric(vec)
    
    #return vector
    return(vec)
}

world.data$Density.P.Km2 <- remove.char(world.data$Density.P.Km2)
world.data$Agricultural.Land.Percent <- remove.char(world.data$Agricultural.Land.Percent)
world.data$Land.Area.Km2 <- remove.char(world.data$Land.Area.Km2)
world.data$Armed.Forces.size <- remove.char(world.data$Armed.Forces.size)
world.data$Co2.Emissions <- remove.char(world.data$Co2.Emissions)
world.data$CPI <- remove.char(world.data$CPI)
world.data$CPI.Change <- remove.char(world.data$CPI.Change)
world.data$Forested.Area.Percent <- remove.char(world.data$Forested.Area.Percent)
world.data$Gasoline.Price <- remove.char(world.data$Gasoline.Price)
world.data$GDP <- remove.char(world.data$GDP)
world.data$Gross.Primary.Education.Enrollment.Percent <- remove.char(world.data$Gross.Primary.Education.Enrollment.Percent)
world.data$Gross.Tertiary.Education.Enrollment.Percent <- remove.char(world.data$Gross.Tertiary.Education.Enrollment.Percent)
world.data$Minimum.wage <- remove.char(world.data$Minimum.wage)
world.data$Out.of.pocket.health.expenditure <- remove.char(world.data$Out.of.pocket.health.expenditure)
world.data$Population <- remove.char(world.data$Population)
world.data$Population.Labor.force.participation.Percent <- remove.char(world.data$Population.Labor.force.participation.Percent)
world.data$Tax.Revenue.Percent <- remove.char(world.data$Tax.Revenue.Percent)
world.data$Total.tax.rate <- remove.char(world.data$Total.tax.rate)
world.data$Unemployment.rate <- remove.char(world.data$Unemployment.rate)
world.data$Urban_population <- remove.char(world.data$Urban_population)


# Join the two datasets
final.data <- happy.data %>%
                left_join(world.data, by = join_by(Country == Country.join)) %>%
                select(-Country.y) %>%
                dplyr::filter(!(Country %in% c('Kosovo', 'Hong Kong S.A.R. of China', 'State of Palestine', 'Taiwan Province of China', 'State of Palestine'))) %>%
                mutate(iso.code = countrycode::countrycode(Country, origin = 'country.name', destination = 'iso3c'))

# Write the text for the plot into the dataset as a variable
# This will be called in the plot later
final.data$hover <- with(final.data, paste(Country, '<br>', 'Population (in millions):', round(Population / 1000000,2), '<br>', 
                          'Life.expectancy: ', Life.expectancy, '<br>', 'GDP Per Cap: $', round(exp(Logged.GDP.per.capita),0)))

```

## Guiding Question

- What factors in a country lead to happiness? 
- Our group decided to look at factors such as 
  - Geography
  - Economics 
  - Health
  
## The Data

- The happiness data was pulled from the World Happiness Report 2023 (Source)[https://worldhappiness.report/data/]
- Additional country statistics were merged to the dataset from the following (Kaggle dataset)[https://www.kaggle.com/datasets/nelgiriyewithana/countries-of-the-world-2023]
- Finally the `countrycode` package was used to import the continent and the iso country code
- The dataset contains 133 countries
  - Kosovo, Hong Kong, State of Palestine, Taiwan, and Palestine were dropped because from the dataset because the country statistics dataset did not have values for them
  - The World Happiness Report did not survey every country in the world, so we **do not** have all countries

## Map of Happiness

```{r, out.width='100%'}
# Load the datasets
world.data <- read.csv('world-data-2023.csv')
happy.data <- read.csv('WHR2023.csv')

# Select and rename columns in happy.data
happy.data <- happy.data %>%
  select(Country.name, Ladder.score) %>%
  rename(Country = Country.name, happiness.score = Ladder.score)

# Prepare world data: you might need to adjust the columns based on your actual world-data-2023.csv structure
# This example assumes minimal preprocessing is needed

# Convert country names in world data to ensure they match those in happy.data
# Use countrycode or manual renaming as needed (shown in your original script)
# This step is simplified for brevity; please adjust according to your actual data

# Join datasets based on country names
# Ensure country names match between datasets; this example assumes they do
final.data <- left_join(happy.data, world.data, by = "Country")

# Add iso codes for plotting
final.data$iso.code <- countrycode(final.data$Country, origin = 'country.name', destination = 'iso3c')

# Generate hover text for plot
final.data$hover <- with(final.data, paste(Country, '<br>', 'Happiness Score:', round(happiness.score, 2)))

# Plotting
plot_ly(data = final.data, type = 'choropleth', locations = ~iso.code, z = ~happiness.score, 
        text = ~hover, colors = 'Spectral') %>%
  colorbar(title = 'Happiness Score') %>%
  layout(title = '2023 World Happiness Score for 133 Countries')

```

- The Western World appears to have the most happiness
- In particular the Nordic Countries (Finland, Norway, Sweden, and Denmark) have the most happiness
- African countries are the most unhappy, followed by countries in the middle east
- Surprisingly Ukraine is not incredibly unhappy


