---
title: "Analyzing the Worlds Happiness"
author: "Owen Gallagher, Jia Qi Lin, Jiacheng Wang"
date: "`r Sys.Date()`"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(plotly)
library(countrycode)

#load in world data
final.data <- read.csv('HW5.data.csv')
final.data$iso.code <- countrycode(final.data$Country, "country.name", "iso3c")

```

## Guiding Question

- What factors in a country lead to happiness? 
- Our group decided to look at factors such as 
  - Geography
  - Economics 
  - Health
  
## The Data

- The happiness data was pulled from the World Happiness Report 2023 (Source)[https://worldhappiness.report/data/]
- Additional country statistics were merged to the dataset from the following (Kaggle dataset)[https://www.kaggle.com/datasets/nelgiriyewithana/countries-of-the-world-2023]
- Finally the `countrycode` package was used to import the continent and the iso country code
- The dataset contains 133 countries
  - Kosovo, Hong Kong, State of Palestine, Taiwan, and Palestine were dropped because from the dataset because the country statistics dataset did not have values for them
  - The World Happiness Report did not survey every country in the world, so we **do not** have all countries

## Map of Happiness

```{r, out.width='100%'}
plot_ly(data = final.data, type = 'choropleth', locations = ~iso.code, z = ~happiness.score, 
        text = ~hover, colors = 'Spectral') %>%
  colorbar(title = 'Happiness Score') %>%
  layout(title = '2023 World Happiness Score for 133 Countries')
```

- The Western World appears to have the most happiness
- In particular the Nordic Countries (Finland, Norway, Sweden, and Denmark) have the most happiness
- African countries are the most unhappy, followed by countries in the middle east
- Surprisingly Ukraine is not incredibly unhappy


