---
title: "Visualizing Japan's Shinkansen (Bullet Train) System"
author: "Minsu Kim, Paul McDonnell, Jiacheng Wang"
date: "2024-04-24"
output: 
  ioslides_presentation:
    widescreen: yes
    css: styles.css
runtime: shiny
---


```{r setup, include=FALSE, echo=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(dplyr)
library(geosphere)
library(igraph)
library(RColorBrewer)
library(plotly)
library(sf)
library(stringr)
library(tidyverse)
library(tmaptools)
library(tmap)

Sys.setenv("MAPBOX_TOKEN" = "pk.eyJ1IjoicG1jZG9ubmVsbCIsImEiOiJjbHV1NXMyeXgwNmU5Mm1wNmVpNXJuZHV5In0.9Re3QXfhIw5NpkLsF-gr1Q")

data <- read.csv("shinkansen_stations_geocoded.csv")
tracks <- st_read("Japan - Shinkansen Lines.kml")
```


## Guiding Questions
- How has the Shinkansen system expanded since the construction of the earliest stations?
- How are Shinkansen stations (and lines) distributed geographically across Japan in 2024?
- What are the key characteristics of Shinkansen stations in Japan?
- What are the preferences of operating companies regarding Shinkansen lines and distances from Tokyo Station?
- What is the relationship between a station's year of opening and the distance from Tokyo Station?
  
  
## Importance 
- The Shinkansen system is a key component of Japan's economy and culture, connecting major cities and regions.
- Understanding the system's development and characteristics can provide insights into Japan's transportation infrastructure.
- Analyzing the relationship between station attributes and operating companies can inform future expansion and improvements.
- Visualizing the Shinkansen system can enhance public awareness and appreciation of this iconic transportation network.


## Data
- The data for the 113 Shinkansen stations were retrieved from a Kaggle dataset (https://www.kaggle.com/datasets/japandata509/shinkansen-stations-in-japan/data).
- Location information was added throgh Google API using the `geocode` function from the `ggmap` package, achieving 97 successful matches.
- 8 locations were matched and added from the Japan Cities Database at simplemaps.com (https://simplemaps.com/data/jp-cities).
- 8 locations were manually hardcoded.


## Variables 
- Station_Name: the name of the station (English)
- Shinkansen_Line: the line that the station belongs
- Year: The year when the station opened
- Prefecture: The prefecture where the station is located
- Distance.from.Tokyo.st: the distance (km) from Tokyo station to the station
- Company: the company that has the station
- lat: latitude 
- lon: longitude


<!-- Paul's Slides -->
## Expansion of the Shinkansen System 
```{r, echo=FALSE, warning=FALSE, message=FALSE}
years <- sort(unique(data$Year))
data$Company <- as.factor(data$Company)
color_palette <- brewer.pal(length(levels(data$Company)), "Set2")
company_colors <- setNames(color_palette, levels(data$Company))
frames <- lapply(years, function(year) {
  frame_data <- data %>% filter(Year <= year)
  list(
    data = list(
      list(
        type = 'scattermapbox',
        lon = frame_data$lon,
        lat = frame_data$lat,
        text = frame_data$Station_Name,
        mode = 'markers',
        marker = list(size = 6, color = company_colors[frame_data$Company]),
        showlegend = FALSE  
      )
    ),
    name = as.character(year)
  )
})
initial_year <- min(years)
initial_data <- data %>% filter(Year <= initial_year)
plot2 <- plot_mapbox() %>%
  layout(
    mapbox = list(
      style = "mapbox://styles/mapbox/light-v10",
      accesstoken = Sys.getenv("MAPBOX_TOKEN"),
      center = list(lon = 138.849, lat = 36.3),
      zoom = 5
    ),
    legend = list(orientation = 'v', y = 1, x = 1, traceorder = 'normal'),
    width = 700,    
    height = 500
  )
plot2 <- plot2 %>%
  add_trace(
    data = initial_data,
    type = 'scattermapbox',
    lon = ~lon,
    lat = ~lat,
    text = ~Station_Name,
    mode = 'markers',
    marker = list(size = 6, color = company_colors[initial_data$Company]),
    showlegend = FALSE
  )
for (company in levels(data$Company)) {
  plot2 <- plot2 %>%
    add_trace(
      type = 'scattermapbox',
      lon = NA,  
      lat = NA,
      mode = 'markers',
      marker = list(size = 10, color = company_colors[company]),
      name = company,
      showlegend = TRUE 
    )
}
plot2$x$frames <- frames
plot2 <- plot2 %>%
  layout(
    sliders = list(
      list(
        active = 0,
        steps = lapply(years, function(y) {
          list(
            method = "animate",
            args = list(list(y), list(mode = "immediate", frame = list(duration = 300))),
            label = y
          )
        }),
        x = 0.1,
        len = 0.9,
        xanchor = "left",
        yanchor = "top"
      )
    )
  )
plot2
```


## Shinkansen Line Data
- The kaggle dataset described did not contain data for the various Shinkansen *lines* throughout Japan.
- The coordinate sets for 8 Shinkansen Lines were obtained via a .KML file, downloadable through Google Maps (https://www.google.com/maps/d/viewer?mid=1lVfyVEXxwWxrMjp9x1-9IDWLGlg&hl=en&ll=36.43907144161917%2C139.30170312654266&z=6).
- Hardcoding of lines not included in this file was less feasible (linestrings; not long/lat coordinate pairs)


## Map of the Present Day Shinkansen System (Almost!)
```{r, echo=FALSE, warning=FALSE, message=FALSE}
tohoku_line <- tracks[tracks$Name == "Tohoku Shinkansen", ]
akita_line <- tracks[tracks$Name == "Akita Shinkansen", ]
yamagata_line <- tracks[tracks$Name == "Yamagata Shinkansen", ]
joetsu_line <- tracks[tracks$Name == "Joetsu Shinkansen", ]
tokaido_line <- tracks[tracks$Name == "Tokaido Shinkansen", ]
sanyo_line <- tracks[tracks$Name == "Sanyo Shinkansen", ]
kyushu_line <- tracks[tracks$Name == "Kyushu Shinkansen", ]
nagano_line <- tracks[tracks$Name == "Nagano Shinkansen", ]
plot_mapbox() %>%
  add_markers(data = data, x = ~lon, y = ~lat, color = ~Company, hoverinfo = "text",
              text = ~paste0("<b>Station_Name:", Station_Name,"<br><b>Company: </b>", 
                             Company, "<br><b>Year: </b>", Year)) %>%
  add_sf(data = tohoku_line, color = I("orange"), name = "Tohoku Line", inherit = FALSE) %>%
  add_sf(data = akita_line, color = I("red"), name = "Akita Line", inherit = FALSE) %>%
  add_sf(data = yamagata_line, color = I("gold"), name = "Yamagata Line", inherit = FALSE) %>%
  add_sf(data = joetsu_line , color = I("plum"), name = "Joetsu Line", inherit = FALSE) %>%
  add_sf(data = tokaido_line, color = I("turquoise"), name = "Tokaido Line", inherit = FALSE) %>%
  add_sf(data = sanyo_line, color = I("green"), name = "Sanyo Line", inherit = FALSE) %>%
  add_sf(data = kyushu_line, color = I("purple"), name = "Kyushu Line", inherit = FALSE) %>%
  add_sf(data = nagano_line, color = I("pink"), name = "Nagano Line", inherit = FALSE) %>%
  layout(title = "The Shinkansen System",
         xaxis = list(title = ""),
         yaxis = list(title = ""),
         margin = list(l = 25, r = 25, b = 25, t = 50),
         font = list(family = "Arial", color = "black"),
         mapbox = list(style = "mapbox://styles/mapbox/light-v10",
                       accesstoken = Sys.getenv("MAPBOX_TOKEN"),
                       center = list(lon = 138.849, lat = 36.3),
                       zoom = 5))
```


<!-- Minsu's Slides -->
## Operating Companies - Number of Stations
```{r, echo=FALSE, warning=FALSE, message=FALSE}
dat <- data %>%
  mutate(Line_name = strsplit(Shinkansen_Line, ",")) %>%
  tidyr::unnest(Line_name) %>%
  select(-Shinkansen_Line) %>%
  rename(Shinkansen_Line = Line_name)
dat$Shinkansen_Line <- gsub("-", "_", dat$Shinkansen_Line)
dat$Shinkansen_Line <- gsub("Shinknsen", "Shinkansen", dat$Shinkansen_Line)
cnt_by_company <- dat %>%
  count(Company, Shinkansen_Line) %>%
  group_by(Company) %>%
  mutate(total_n = sum(n)) %>%
  ungroup()
annotations <- list()
for (i in 1:length(unique(cnt_by_company$Company))) {
  annotations[[i]] <- list(
    x = unique(cnt_by_company$Company)[[i]],
    y = unique(cnt_by_company$total_n)[[i]],
    text = unique(cnt_by_company$total_n)[[i]],
    yanchor = "bottom",
    showarrow = FALSE
  )
}
cnt_by_company %>%
  plot_ly(
    x = ~Company,
    y = ~n,
    color = ~Shinkansen_Line,
    colors = "Spectral",
    hoverinfo = "text",
    text = ~ paste(n, "stations in", Shinkansen_Line)
  ) %>%
  add_bars() %>%
  layout(
    barmode = "stack",
    xaxis = list(title = "Company"),
    yaxis = list(title = "Number of Shinkansen Stations"),
    title = "Number of Shinkansen Stations by Company",
    annotations = annotations
  )
```
<span style="font-size: 0.65em;">
Note: There are 7 intersection stations, and we count the number of `Shinkansen_Line` twice for the intersection stations. 
For this, we split the string in the `Shinkansen_Line` column when it has both Shinkansen line names and unnest the data.
</span>


## Operating Companies - Distance
```{r, echo=FALSE, warning=FALSE, message=FALSE}
colramp <- colorRamp(c("darkblue", "forestgreen", "red", "darkred", "orange"))
boxp_company <- dat %>%
  plot_ly(
    y = ~Distance.from.Tokyo.st,
    color = ~Company,
    colors = colramp,
    type = "box",
    showlegend = FALSE
  ) %>%
  layout(
    xaxis = list(title = "Company"),
    title = "Boxplots of Distance from Tokyo Station by Company"
  )
boxp_line <- dat %>%
  mutate(Line = str_replace_all(Shinkansen_Line, "_Shinkansen", "")) %>%
  plot_ly(
    y = ~Distance.from.Tokyo.st,
    color = ~Line,
    type = "box",
    showlegend = FALSE
  ) %>%
  layout(
    xaxis = list(title = "Shinkansen Line"),
    title = "Boxplots of Distance from Tokyo Station by Shinkansen Line"
  )
subplot(boxp_company, boxp_line, nrows = 1, titleX = TRUE, margin = 0.03) %>%
  layout(
    yaxis = list(title = "Distance from Tokyo Station (km)"),
    title = 
      "Boxplots of Distance from Tokyo Station by Company and Shinkansen Line",
    plot_bgcolor = "#e5ecf6"
    )
```


<!-- Jiacheng's Slides -->
## Network Plot of the Shinkansen System
```{r, echo=FALSE, warning=FALSE, message=FALSE}
nodes <- unique(c("Tokyo", data$Station_Name))
edges <- data.frame(from = rep("Tokyo", length(nodes)-1), to = nodes[nodes != "Tokyo"]) 
coords <- data.frame(name = nodes, lat = data$lat[match(nodes, data$Station_Name)], 
                     lon = data$lon[match(nodes, data$Station_Name)])

years <- data$Year
distances_to_tokyo <- data$Distance.from.Tokyo.st[match(nodes, data$Station_Name)]
g <- graph_from_data_frame(edges, directed = TRUE)

distances <- mapply(function(from, to) {
  if (from != to) {
    from_coords <- c(coords$lon[coords$name == from], coords$lat[coords$name == from])
    to_coords <- c(coords$lon[coords$name == to], coords$lat[coords$name == to])
    return(geosphere::distHaversine(from_coords, to_coords))
  } else {
    return(0) 
  }
}, edges$from, edges$to)
E(g)$weight <- distances

layout <- layout_with_kk(g)
x <- layout[, 1]
y <- layout[, 2]
names <- V(g)$name

edge_x <- c()
edge_y <- c()
for (eid in E(g)) {
  edge_x <- c(edge_x, x[head_of(g, eid)], x[tail_of(g, eid)], NA)
  edge_y <- c(edge_y, y[head_of(g, eid)], y[tail_of(g, eid)], NA)
}

plot_ly() %>%
  add_trace(x = ~edge_x, y = ~edge_y, type = 'scatter', mode = 'lines', 
            line = list(color = 'grey', width = 1)) %>%
  add_trace(x = ~x, y = ~y, text = ~names, type = 'scatter', mode = 'markers',
            marker = list(size = 10, color = ~years, coloraxis = 'coloraxis'),
            textposition = 'bottom center', hoverinfo = "text", 
            hovertemplate = "<b>%{text}</b><br>Year Opened: %{marker.color}<br>Distance to Tokyo: %{customdata} km<extra></extra>",
            customdata = ~distances_to_tokyo) %>%
  layout(xaxis = list(title = "", showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(title = "", showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         showlegend = FALSE,
         coloraxis = list(colorscale = 'Viridis', colorbar = list(title = "Year Opened"))) 
```
<div style="font-size: 0.8em;">
  <ul>
    <li>The length of the edges represents the distance between the stations to Tokyo.</li>
    <li>The color of the nodes represents the year the station was opened.</li>
  </ul>
</div>


## Summary
- Retrieve and clean the Shinkansen station data.
- Retrieve the Shinkansen line data and define lines.  
- Map plots - Expansion of Shinkansen System up to the present day.
- Box plots - Analysing characteristics of the operating companies.
- Network plot - Exploring the relationship between year of opening and the distance to Tokyo.


