---
title: "Is the age of electric vehicles coming?"
author: "Kuanjung Huang, Jiacheng Wang, Ping Yang"
date: "2024-05-15"
output: 
  ioslides_presentation:
    widescreen: yes
#runtime: shiny
---


```{r setup, include=FALSE, echo=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(dplyr)
library(RColorBrewer)
library(plotly)
library(reshape2)
```


## Guiding Questions
- Whether the fuel price is a key factor in the development of electric vehicles?
- How do fuel taxes affect the sales of hybrid electric vehicles?
- What are the trends in hybrid electric vehicle sales by different manufacturers?


## Data
- [Average Retail Fuel Prices in the United States](https://afdc.energy.gov/fuels/prices.html)

- [Fuel Taxes by Country](https://www.compareyourcountry.org/taxing-energy?cr=oecd&lg=en&page=2&visited=1)

- [U.S. HEV Sales by Model](https://afdc.energy.gov/data/10301)

<!-- Jiacheng's Slides -->
## Average Retail Fuel Prices in the United States
```{r, echo=FALSE, warning=FALSE, message=FALSE}
fuel_prices <- read.csv("gasoline_prices_by_year.csv")
plot_ly(fuel_prices, x = ~Year, 
        y = ~Gasoline_Price, 
        type = 'scatter', 
        mode = 'lines+markers', 
        name = 'Gasoline Price ($/gallon)', 
        marker = list()) %>%
  add_trace(y = ~Gasoline_Price_2023, 
            name = 'Gasoline Price (2023 $/gallon)', 
            marker = list()) %>%
  layout(xaxis = list(title = 'Year'),
         yaxis = list(title = 'Gasoline Price ($/gallon)'),
         width = 1000)
```
<span style="font-size: 0.7em;">
**Inflation-adjusted gasoline prices** were calculated based on: $P_{2023} = P \times Inflation factor$.
</span>

## Fuel Taxes by Country
```{r, echo=FALSE, warning=FALSE, message=FALSE}
fuel_tax <- read.csv("fuel_taxes_by_country.csv")
ordered_countries <- fuel_tax$Country[order(-fuel_tax$Gasoline)]
plot_ly(fuel_tax, x = ~Country, y = ~Gasoline, type = 'bar', name = 'Gasoline') %>%
  add_trace(y = ~Diesel, name = 'Diesel') %>%
  layout(
    xaxis = list(
      title = 'Country',
      tickangle = -45,
      tickfont = list(size = 13),
      automargin = TRUE,
      type = 'category',
      categoryorder = 'array',
      categoryarray = ~ordered_countries
    ),
    yaxis = list(title = 'Fuel Tax ($/gallon)'),
    barmode = 'group',
    width = 1000
  )
```

<!-- Ping's Slides -->
## U.S. HEV Sales by Model
```{r, echo=FALSE, warning=FALSE, message=FALSE}
data <- readxl::read_excel("hev_sale.xlsx")
hev_sales_long <- melt(data, id.vars = "Vehicle", variable.name = "year", value.name = "sales")
total_sales <- hev_sales_long %>%
  group_by(Vehicle) %>%
  summarise(total_sales = sum(sales))
hev_sales_long <- merge(hev_sales_long, total_sales, by = "Vehicle")
hev_sales_long$Vehicle <- factor(hev_sales_long$Vehicle, levels = total_sales$Vehicle[order(total_sales$total_sales, decreasing = FALSE)])
num_models <- length(unique(hev_sales_long$Vehicle))
custom_palette <- colorRampPalette(brewer.pal(8, "Set3"))(num_models)

hev_sales_long %>%
  plot_ly() %>%
  add_bars(
    x = ~year, 
    y = ~sales, 
    color = ~Vehicle, 
    colors = custom_palette, 
    text = ~paste("Year: ", year, "<br>Model: ", Vehicle, "<br>Sales: ", sales, "<br>Total Sales: ", total_sales), 
    hoverinfo = "text"
    ) %>%
  layout(
    xaxis = list(title = "Year"),
    yaxis = list(title = "Sales"),
    barmode = "stack",
    showlegend = FALSE
  )
```

## Summary
- Retrieve data from DOE Alternative Fuels Data Center.
- **Line chart** - Trend of alternative and traditional motor fuel prices from 2000 to 2024.
- **Bar chart** - Comparison of gasoline and diesel taxes in different countries.
- **Stacked bar chart** - Trend of sales by hybrid electric vehicle (HEV) model, 1999-2019.
- **Shiny APP** - 


