---
title: "2023 World Happiness Report"
author: "Owen Gallagher, Nikki Lin, Jiacheng Wang"
date: "2024-03-13"
output: 
  ioslides_presentation:
    widescreen: yes
    css: styles.css
runtime: shiny # comment out to generate a static html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Guiding Question
- What factors in a country lead to happiness? 
- Our group decided to look at factors such as 
  - Geography
  - Economics 
  - Health
  
## The Data
- The happiness data was pulled from the World Happiness Report 2023 
[Source](https://worldhappiness.report/data/)

- Additional country statistics were merged to the dataset from the following 
[Kaggle dataset](https://www.kaggle.com/datasets/nelgiriyewithana/countries-of-the-world-2023)

- Finally the `countrycode` package was used to import the continent and the 
iso country code

- The dataset contains 133 countries

  - Kosovo, Hong Kong, State of Palestine, Taiwan, and Palestine were dropped 
  because from the dataset because the country statistics dataset did not have 
  values for them
  
  - The World Happiness Report did not survey every country in the world, so we 
  **do not** have all countries

## 2023 World Happiness Score for 133 Countries {.small-text}
```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(plotly)
library(countrycode)

final.data <- read.csv('HW5.data.csv')
Country <- final.data$Country
iso.code = countrycode::countrycode(Country, origin = 'country.name', 
                                    destination = 'iso3c')
final.data$hover <- with(final.data, paste(Country, '<br>', 
                                           'Population (in millions):', 
                                           round(Population / 1000000,2), 
                                           '<br>', 
                          'Life.expectancy: ', 
                          Life.expectancy, 
                          '<br>', 'GDP Per Cap: $', 
                          round(exp(Logged.GDP.per.capita),0)))

plot_ly(data = final.data, type = 'choropleth', 
        locations = ~iso.code, z = ~happiness.score, 
        text = ~hover, colors = 'Spectral') %>%
  colorbar(title = 'Happiness Score')
```

- The Western World appears to have the most happiness
- The Nordic Countries (Finland, Norway, Sweden, and Denmark) have the most happiness
- African countries are the most unhappy, followed by countries in the middle east
- Surprisingly Ukraine is not incredibly unhappy


## Nikki's Slide 1
```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(readr)
library(stargazer)
library(tidyverse)
library(broom)
library(countrycode)
library(plotly)
library(patchwork)
library(stargazer)

hw5 <- read.csv("HW5.data.csv")

psydata <- hw5 %>% 
  dplyr::select(Country, happiness.score, Social.support, 
                Freedom.to.make.life.choices, Perceptions.of.corruption) 

psydata2 <- psydata %>% 
    pivot_longer(c("Social.support", "Freedom.to.make.life.choices", 
                   "Perceptions.of.corruption"), 
                 names_to = "Predictor", values_to = "Score")

psydata2$Continent <- countrycode(sourcevar = psydata2[["Country"]], 
                                  origin = "country.name", 
                                  destination = "continent")

psy_pred_model2 <- lm(happiness.score ~ Social.support + 
                        Freedom.to.make.life.choices + 
                        Perceptions.of.corruption, data = hw5)

plot1 <- ggplot(psydata2, aes(Score, happiness.score, color = Continent)) + 
  geom_point(aes(ids = Country), alpha = 0.5) +
  theme(legend.title = element_blank()) +
  facet_wrap(~Predictor, labeller = 
               labeller(Predictor = 
                          c("Freedom.to.make.life.choices" = 
                              "Life Choice Freedom", 
                            "Perceptions.of.corruption" = 
                              "Corruption Perception",
                            "Social.support" = 
                              "Social Support"))) + 
  geom_smooth(method='lm', se = F, alpha = 0.5) + 
  theme_bw() + 
  ylab("Happiness Score")

plotly1 <- ggplotly(plot1)

plotly1
```

## Nikki's Slide 2
```{r, echo=FALSE, warning=FALSE, message=FALSE}
plot2 <- ggplot(psydata2, aes(x = Score, y = happiness.score, 
                              color = Predictor)) + 
  geom_smooth(method='lm', aes(fill = Predictor), se = TRUE, alpha = 0.5) + 
  scale_fill_manual(values=c(Freedom.to.make.life.choices="plum3",
                             Perceptions.of.corruption="khaki1",
                             Social.support= "mistyrose")) +
    scale_color_manual(values=c(Freedom.to.make.life.choices="mediumpurple4",
                             Perceptions.of.corruption="orange2",
                             Social.support="palevioletred3")) + 
  theme_bw() +
  ylab("Happiness Score")

plotly2 <- ggplotly(plot2, tooltip = c("text"))

plotly2
```

## Multiple Regression Predicting Happiness Score
```{r, echo=F, results='asis'}
stargazer(psy_pred_model2, header=F, type = "html", 
          covariate.labels = c("Social support", 
                               "Life choice freedom", 
                               "Corruption perception",
                               "Constant",
          dep.var.labels   = ("Happiness score")))
```

<!--## Economic Plot with GDP and Happiness Score
```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
library(plotly)

# Data preparation and plot GDP with Happiness Score
hw_data <- read.csv("HW5.data.csv", stringsAsFactors = TRUE)
countries_all <- read.csv("region.csv", stringsAsFactors = TRUE)
combined_data <- left_join(hw_data, countries_all, by = "Country")
write.csv(combined_data, "HW5_with_region.csv")
happiness2023 <- read.csv("HW5_with_region.csv", stringsAsFactors = TRUE)

# Plot GDP with Happiness Score
gg <- ggplot(happiness2023, aes(x = Logged.GDP.per.capita, 
                                y = happiness.score, 
                                color = Region)) + 
  geom_point(aes(size = Population), show.legend = 'size') + 
  scale_x_log10() + 
  theme(legend.title = element_blank()) +
  guides(color = guide_legend("Region"), size = FALSE)

ggplotly(gg, tooltip = c("x", "y", "color", "size")) 
```
-->

## Shiny APP
```{r shiny, echo=FALSE, warning=FALSE, message=FALSE}
library(shiny)
library(markdown)
library(ggplot2)
library(dplyr)
library(plotly)
library(shinythemes)


# Data preparation
happiness2023 <- read.csv("HW5_with_region.csv", stringsAsFactors = TRUE)
colnames <- colnames(happiness2023) %>% .[-c(1, 2, 3)] %>% .[-length(.)]

# Define UI 
ui <- navbarPage(
  theme = shinytheme("journal"),
  title = "World Happiness",
  tabPanel("Plot",
           sidebarPanel(
             width = 3,
             uiOutput("variable_1"),
             uiOutput("variable_2"),
             helpText("ctrl/shift for multiple choice"),
             uiOutput("country"),
             hr(),
             span("Data source:", 
                  tags$a("WHR 2023",
                         href = "http://worldhappiness.report"))),
           mainPanel(plotlyOutput("scatterplot"))),
           
  
  tabPanel("About", includeMarkdown("About.md")))

      
# Define server 
server <- function(input, output, session) {
  
  output$variable_1 <- renderUI({
    selectInput("variable_1", "Y-variable:", 
                choices = colnames, selected = "happiness.score")})
  
  output$variable_2 <- renderUI({
    selectInput("variable_2", "X-variable:", 
                choices = colnames, selected = "happiness.score")})

  data_2023 <- reactiveValues(data = happiness2023)
  
  filtered_data <- reactive({
    data_2023$data %>% 
      filter(Country %in% c(input$country)) %>% 
      arrange(Country)
    })
  
  countries <- reactive({ 
    df_small <- 
      happiness2023 %>% 
      select(Country) %>% 
      droplevels()
    c("Select All", levels(df_small$Country))
    })
  
  output$country <- renderUI ({
    selectInput("country", "Countries:", 
                multiple = TRUE,
                choices = countries(), 
                selected = "Select All",
                selectize = FALSE,
                size = 10)
  })

  observe({
    if ("Select All" %in% input$country) {
      selected_choices <- setdiff(countries(), "Select All")
      updateSelectInput(session, "country", selected = selected_choices)
    }
  })

  output$scatterplot <- renderPlotly({
    if (is.null(data_2023$data)) return()
    
    p <- ggplot(filtered_data()) +
      geom_point(aes_string(x = input$variable_2, 
                            y = input$variable_1, 
                            colour = "Region", 
                            label = "Country"), 
                 size = 3) +
      ggtitle(paste0(input$variable_1, " vs. ", input$variable_2))
  })
}

# Run the application 
shinyApp(ui = ui, server = server)
```


## Summary

- 
- 
- Created a Shiny app to visualize the relationship between happiness score and other variables
