---
title: "Is the age of electric vehicles coming?"
author: "Kuanjung Huang, Jiacheng Wang, Ping Yang"
date: "2024-05-15"
output: 
  ioslides_presentation:
    widescreen: yes
#runtime: shiny
---


```{r setup, include=FALSE, echo=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(dplyr)
library(RColorBrewer)
library(plotly)
library(reshape2)
```


## Guiding Questions
- Whether the fuel price is a key factor in the development of electric vehicles (EV)?
- How do fuel taxes affect the sales of EV?
- What are the trends in EV sales by different manufacturers?
- How fast-growing is the EV business?
- Are there more people driving EVs in a more wealthy state?


## Data
- DOE Alternative Fuels Data Center ([https://afdc.energy.gov/data](https://afdc.energy.gov/data))
  - Average retail fuel prices in the US
  - Fuel taxes by country
  - EV sales in the US by model
  - Number of vehicles registrations by state and by energy type
- International Energy Agency ([https://www.iea.org/data-and-statistics](https://www.iea.org/data-and-statistics))
  - Electric car registrations in United States and Europe
- United Census Bureau ([link](https://data.census.gov/map/040XX00US51,50,55,56,12,53,54,10,15,16,13,19,17,18,40,41,44,45,01,42,48,49,05,02,46,47,08,09,06,30,72,33,34,31,32,37,38,35,36,39,22,23,20,21,26,27,24,25,28,29,04/ACSST1Y2018/S1901/S1901_C01_001E?q=S1901:%20INCOME%20IN%20THE%20PAST%2012%20MONTHS%20(IN%202018%20INFLATION-ADJUSTED%20DOLLARS)&layer=VT_2018_040_00_PP_D1&loc=34.3659,-21.2851,z0.3809))
  - US median income by state 


<!-- Jiacheng's Slides -->
## Average Retail Fuel Prices in the US
```{r, echo=FALSE, warning=FALSE, message=FALSE}
fuel_prices <- read.csv("gasoline_prices_by_year.csv")
plot_ly(fuel_prices, x = ~Year, 
        y = ~Gasoline_Price, 
        type = 'scatter', 
        mode = 'lines+markers', 
        name = 'Gasoline Price ($/gallon)', 
        marker = list()) %>%
  add_trace(y = ~Gasoline_Price_2023, 
            name = 'Gasoline Price (2023 $/gallon)', 
            marker = list()) %>%
  layout(xaxis = list(title = 'Year'),
         yaxis = list(title = 'Gasoline Price ($/gallon)'),
         width = 900)
```
<span style="font-size: 0.7em;">
**Inflation-adjusted gasoline prices** were calculated based on: $P_{2023} = P \times Inflation factor$.
</span>


## Fuel Taxes by Country in 2018
```{r, echo=FALSE, warning=FALSE, message=FALSE}
fuel_tax <- read.csv("fuel_taxes_by_country.csv")
ordered_countries <- fuel_tax$Country[order(-fuel_tax$Gasoline)]
plot_ly(fuel_tax, x = ~Country, y = ~Gasoline, type = 'bar', name = 'Gasoline') %>%
  add_trace(y = ~Diesel, name = 'Diesel') %>%
  layout(
    xaxis = list(
      title = 'Country',
      tickangle = -45,
      tickfont = list(size = 13),
      automargin = TRUE,
      type = 'category',
      categoryorder = 'array',
      categoryarray = ~ordered_countries
    ),
    yaxis = list(title = 'Fuel Tax ($/gallon)'),
    barmode = 'group',
    width = 900
  )
```
<ul style="font-size: 0.7em;">
  <li>Switzerland, Iceland, Mexico, and the US are the only countries with higher taxes on diesel than gasoline.</li>
  <li>In general, European countries tend to tax petroleum fuels at the highest level globally.</li>
</ul>


<!-- Ping's Slides -->
## EV Registrations in Europe and the US
```{r, echo=FALSE, warning=FALSE, message=FALSE}
ev_registrations <- read.csv("ev_registration.csv")
plot_ly(ev_registrations, x = ~Year, y = ~Europe, type = 'bar', name = 'Europe') %>%
  add_trace(y = ~US, name = 'United States') %>%
  layout(
    xaxis = list(
      title = 'Year',
      tickfont = list(),
      automargin = TRUE
    ),
    yaxis = list(title = 'EV Registrations (million)'),
    barmode = 'group',
    width = 900
  )
```


## The US EV Sales by Model
```{r, echo=FALSE, warning=FALSE, message=FALSE}
data <- readxl::read_excel("10567_pev_sales_2-28-20.xlsx")
hev_sales_long <- melt(data, id.vars = "Vehicle", variable.name = "year", value.name = "sales")
total_sales <- hev_sales_long %>%
  group_by(Vehicle) %>%
  summarise(total_sales = sum(sales))
hev_sales_long <- merge(hev_sales_long, total_sales, by = "Vehicle")
hev_sales_long$Vehicle <- factor(hev_sales_long$Vehicle, levels = total_sales$Vehicle[order(total_sales$total_sales, decreasing = FALSE)])
num_models <- length(unique(hev_sales_long$Vehicle))
custom_palette <- colorRampPalette(brewer.pal(11, "Set3"))(num_models)
hev_sales_long %>%
  plot_ly() %>%
  add_bars(
    x = ~year, 
    y = ~sales, 
    color = ~Vehicle, 
    colors = custom_palette, 
    text = ~paste("Year:", year, "<br>Model:", Vehicle, "<br>Sales:", sales, "<br>Total Sales:", total_sales), 
    hoverinfo = "text",
    textposition = "none"
    ) %>%
  layout(
    xaxis = list(title = "Year"),
    yaxis = list(title = "Sales"),
    barmode = "stack",
    showlegend = TRUE,
    width = 900
  )
```


## Summary
- Retrieve data from DOE Alternative Fuels Data Center.
- **Line chart** - Fuel prices could be a factor in the EV market.
- **Bar chart** - Taxes do affect the EV market.
- **Stacked bar chart** - Tesla is the leading EV manufacturer in the US.
- **Shiny APP map + Scatter / Line chart** - EV business is rising but sill extremely small compared to gas cars. The growths are most notable in wealthier states.

